---
title: "Bianca's Notes for CH 8"
author: "Bianca Luedeker"
date: "2024-02-27"
output: pdf_document
---

Load packages

## Vectors

a. Using seq count by 4
```{r}
aa <- seq(100, 300, 5)
aa
```

b. Make a vector of length 100 where the numbers are evenly spaced.

```{r}
seq(676, 942, length.out=100)
```

c. The negative notation.

```{r}
aa[c(-2, -8, -10)]

```


## Matrices

a. Construct a matrix with 6 rows and 4 columns.
```{r}
A <- matrix(1:24, nrow=6, ncol=4, byrow = TRUE)
A
```

b. cbind 
```{r}
C1 <- c(8, 7, 10)
C2 <- c( 11, 5, -6)
C3 <- c( 20, 27, 10)

cbind(C1, C2, C3)
```

c. rbind
```{r}
rbind(C1, C2, C3)
```

d. Extract an element
```{r}
A[2, 3]
```

e. Extract a row
```{r}
A[2,]
```

f. Extract a column.
```{r}
A[,2]
```


## Data Frames

a. Construct a dataframe.
```{r}
my.df <- data.frame(height = c(60, 65, 68),
                    weight = c(125, 119, 140))
my.df
```

b. Extract a column.
```{r}
my.df$height
my.df[,1]
```

c. Extract a value.
```{r}
my.df[2,2]
```


## Lists
a. Construct a list of unlike objects.
```{r}
my.list <- list(vec = aa, matrix = A)
my.list
```

b. Extract using dollar sign notation.

```{r}
my.list$vec
my.list$matrix
```

c. Extract using single brackets.
```{r}
my.list[1]
str(my.list[1])
```

d. Extract using double brackets.  What is the difference?
```{r}
my.list[[1]]
str(my.list[[1]])
```
e. lm is a list.  Show what is in lm.
```{r}
data("iris")
mod <- lm(data=iris, Petal.Length~Petal.Width)
str(mod)
mod$coefficients
```

