---
title: "STA445_CH5_Code"
author: "Levi Mault"
date: "2024-02-20"
output: pdf_document
---

## This is a document that students are filling out with the instructor in class for chapter 5.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Use the mtrcars dataset that is in base R.

```{r}
my.cars <- mtcars
my.cars$cyl <- as.factor(my.cars$cyl)

class(my.cars$cyl)
```

## Run a t-test that compares the mean fuel economy for cars with 6-cylinders 
to 8-cylinders

```{r}
my.cars.1 <- my.cars %>%
  select(mpg, cyl) %>%
  filter( cyl == "6" | cyl == "8")

slice_sample(my.cars.1, n=5) # This shows 5 random rows from the data set
```

## Do the t-test.
```{r}
t.test(data=my.cars.1, mpg ~ cyl, conf.level=0.9)
```

## Note: Disregard most of the code found in ch 5 of the book...

## How does horsepower affect fuel economy?
```{r}
ggplot(data=my.cars, aes(x=hp, y=mpg)) +
  geom_point()
```

```{r}
mpg.mod <- lm(data=my.cars, mpg ~ hp)
summary(mpg.mod)
```

```{r}
predict(mpg.mod, interval="confidence")
```

```{r}
my.cars.w.pred <- cbind(my.cars, predict(mpg.mod, interval="confidence"))

my.cars.w.pred
```


#Plot the points with the fitted regression line and confidence bands
```{r}
ggplot(data=my.cars.w.pred, aes(x=hp, y=mpg)) +
  geom_point() +
  geom_line(aes(y=fit)) + 
  geom_ribbon((aes(ymin=lwr, ymax=upr)), alpha=0.5, fill = "skyblue") # Alpha ~ Transparency
```

